Manuscript 1 code

### 2025-03-06 manuscript_1_code.Rmd
Corrected main figure aesthetics

model_themes <-   theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend

@@

```{r figure 3a plot}
(figure_3a <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.5) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "", title = "") +
  model_themes 
         )
# ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)

@@

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro,
             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_cover_effects_fg, aes(x = functional_richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects_fg, aes(x = functional_richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "", title = "") +
  trends_theme
         )
#ggsave(filename = "output/figure_3b_quad_v2.png", figure_3b, height = 10, width = 14)

@@

## Figure 3 Full Plot
```{r ggpubr fig3, fig.height = 8, fig.width=16}

(figure_3 <- ggarrange(figure_3a, figure_3b, figure_3c, figure_3d, figure_3e, figure_3f, 
                      ncol = 3, nrow = 2, 
                      common.legend = TRUE,
                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
                      legend = "bottom", 
                      font.label = list(size = 26, color = "black", face = "plain")))

@@

### full vs mean plot comparison plot
```{r ggpubr meanfull, fig.height = 8, fig.width=9}

(mean_full_fxn_plot <- ggarrange(figure_3a, figure_3a_2, figure_3b, figure_3b_2, 
                      ncol = 2, nrow = 2, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_3v2.png", figure_3, height = 8, width = 16)
```
```{r figure 3 with mean, fig.height = 8, fig.width=9}
(figure_3_2 <- ggarrange(figure_3a_2, figure_3b_2, figure_3d, figure_3e, 
                      ncol = 2, nrow = 2, 
                      common.legend = TRUE,
                      labels = c("a.", "b.", "c.", "d."),
                      legend = "bottom", 
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_3v3.png", figure_3_2, height = 8, width = 16)

```
# Figure 4

@@

```{r figure 4a}
(figure_4a <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "", title = "") +
  trends_theme
         )

@@

```{r plot figure 4b}
(figure_4b <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_fg, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_fg,
              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "", title = "") +
  trends_theme
         )

@@

```{r figure 4c}
(figure_4c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "", y = "", title = "") +
  trends_theme
         )
# ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)

@@

```{r figure 4d}
(figure_4d <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Taxonomic Richness", title = "") +
  trends_theme
         )

@@

```{r figure 4e}
(figure_4e <- 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_site, 
              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Functional Richness", title = "") +
  trends_theme
         )

@@

```{r figure 4f}
(figure_4f <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "", title = "") +
  trends_theme
         )
# ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
```

## Figure 4 Full Plot
```{r ggpubr fig4, fig.height = 8, fig.width=16}

(figure_4 <- ggarrange(figure_4a, figure_4b, figure_4c, figure_4d, figure_4e, figure_4f, 
                      ncol = 3, nrow = 2, 
                      common.legend = TRUE,
                      heights = c(1,1),
                      widths = c(1,1,1),
                      legend = "bottom", 
                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_4v2.png", figure_4, height = 8, width = 16)
```

@@

(spatial_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Spatial synchrony", title = "") +
  trends_theme
         )

@@

(mean_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = synchrony_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_mean_synch_effects, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_mean_synch_effects, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Site-level Stability\n", x = "Mean plot-level species synchrony", title = "") +
  trends_theme
         )
```

@@

(alpha_gamma_stability_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Plot-level stability", title = "") +
  trends_theme
         )
```

## panel graph
```{r figure 5 panel, fig.height = 6, fig.width=22}
(figure_5 <- ggarrange(mean_synchrony_plot, alpha_gamma_stability_plot, spatial_synchrony_plot,
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_5_option3_v2.png", figure_5, height = 6, width = 22)
```

>>>>

model_themes <-   theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=24),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend

@@

```{r figure 3a plot}
(figure_3a <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.75) + 
  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.5) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Taxonomic richness", title = "") +
  model_themes 
         )
# ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)

@@

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro,
             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.75) + 
  geom_line(data = filtered_cover_effects_fg, aes(x = functional_richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects_fg, aes(x = functional_richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.5) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Functional richness", title = "") +
  model_themes
         )
#ggsave(filename = "output/figure_3b_quad_v2.png", figure_3b, height = 10, width = 14)

@@

## Figure 3 Full Plot
```{r ggpubr fig3 main, fig.height = 8, fig.width=16}

(figure_3 <- ggarrange(figure_3a, figure_3b,  
                      ncol = 2, nrow = 1, 
                      common.legend = TRUE,
                      labels = c("a.", "b."),
                      legend = "bottom", 
                      font.label = list(size = 26, color = "black", face = "plain")) )

@@

```{r figure 3 site supplement, fig.height = 8, fig.width = 16}
(figure_3_supplement <- ggarrange(figure_3d, figure_3e,  
                      ncol = 2, nrow = 1, 
                      common.legend = TRUE,
                      labels = c("a.", "b."),
                      legend = "bottom", 
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_3_site.png", figure_3_supplement, height = 8, width = 16)
```

# Figure 4

@@

```{r figure 4a}
(figure_4a <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Taxonomic richness", title = "") +
  model_themes
         )

@@

```{r plot figure 4b}
(figure_4b <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_fg, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_fg,
              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_x_continuous(breaks = c(0,1,2)) +
  labs(y = "", x = "Functional richness", title = "") +
  model_themes
         )

@@

```{r figure 4c}
(figure_4c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "", title = "") +
  model_themes
         )
# ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)

@@

```{r figure 4d}
(figure_4d <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  labs(y = "Site-level Stability\n", x = "Taxonomic richness", title = "") +
  model_themes
         )

@@

```{r figure 4e}
(figure_4e <- 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_site, 
              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  scale_x_continuous(limits = c(2, 4.2), breaks = c(1,2,3,4)) +
  labs(y = "", x = "Functional richness", title = "") +
  model_themes
         )

@@

```{r figure 4f}
(figure_4f <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "", title = "") +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  model_themes
         )
# ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
```

## Figure 4 Full Plot
```{r ggpubr fig4, fig.height = 6, fig.width=18}

(figure_4 <- ggarrange(figure_4a, figure_4b, figure_4c, 
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      heights = c(1,1),
                      widths = c(1,1,1),
                      legend = "bottom", 
                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_4v3.png", figure_4, height = 6, width = 18)
```

@@

(spatial_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Spatial synchrony", title = "") +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  model_themes
         )

@@

(mean_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = synchrony_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_mean_synch_effects, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_mean_synch_effects, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Site-level Stability\n", x = "Mean plot-level species synchrony", title = "") +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  model_themes
         )
```

@@

(alpha_gamma_stability_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Plot-level stability", title = "") +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  model_themes
         )
```

## panel graph
```{r figure 5 panel, fig.height = 10, fig.width=18}
(figure_5 <- ggarrange(figure_4d, figure_4e, figure_4f, 
                       mean_synchrony_plot, alpha_gamma_stability_plot, spatial_synchrony_plot,
                      ncol = 3, nrow = 2, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_5_v2.png", figure_5, height = 10, width = 18)
```
